---
title: "A system for predicting sports performance"
output: html_document
Author: Gazzella Ingegnosa
---

##Introduction

In this project  we will try to implemente a procedure able to quantify how well people perform excercises.  
We will use data gathered with devices like Fitbit, Jawbon up, etc. from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

##Uploading libraries and dataset

```{r libdataset} 
library(caret)
library(randomForest)
library(YaleToolkit)
library(e1071)

setwd("C:/coursera")
#Upload dataset for project
training <- read.csv("./pml-training.csv")
#Upload dataset for quiz (validation set)
#plmvalidation <- read.table("./pml-testing.csv", sep=",", header=TRUE)

```
##Cross-validation
The procedure consists in splitting in training set into sub-training/test sets
For cross-validation, as we have a large sample we will split dataset in 60% training and 40% test.

```

```{r testtrain}

set.seed(33833)
trainIndex = createDataPartition(training$classe,p=0.6,list=FALSE)
training = training[trainIndex,]
testing = training[-trainIndex,]

```
##Preprocessing to select variables
We explore dataset verifying if there are variables not compiled.
Ploting data shows that there is problably a pattern.

```{r Plotting}

myTitle <-"Sport performance"
qplot(num_window, roll_belt, col=classe,data=training) +  labs(title = myTitle)

```
Dropping null variable.
```{r dropvar}

ListVar <- data.frame(whatis(training, var.name.truncate = 50))
SelectVar<-subset(ListVar, ListVar$missing == 0)
colnames<-as.vector(SelectVar$variable.name)
trainsel<-training[,colnames(training)%in%colnames]


```
How many variables with a low compilation?
We apply near0var procedure to drop variables with a low variability.
Moreover we dropped X, timestamp and user_name traisel2$X.

```{r zerovar}

ListNear0<- nearZeroVar(trainsel,saveMetrics = TRUE)
SelectNear0<-subset(ListNear0, ListNear0$nzv == FALSE)
colnames<-as.vector(row.names(SelectNear0))
trainsel2<-trainsel[,colnames(trainsel)%in%colnames]

trainsel2$cvtd_timestamp <- NULL
trainsel2$user_name <- NULL
trainsel2$X <- NULL


```

##Classification with random forest 
We built a model on training set, then we have evaluated on test set


```{r randfor}

rfFit <- randomForest(trainsel2$classe ~ .,data=trainsel2,importanCE=TRUE, ntree=50)
varImp(rfFit)
head(getTree(rfFit, 2))

```
## Out of sample error
We have repeated and averaged estimated errors

```{r outofsample}

confusionMatrix(trainsel2$classe,predict(rfFit,trainsel2))
confusionMatrix(testing$classe,predict(rfFit,testing))

```







